---
title: "HW9"
author: "Artur Sak"
date: "November 16, 2016"
output: html_document
---
#Part I Los Angeles Daily Weather

```{r}

load("weather2011.rda")
monthNames = c("January", "February", "March", "April",
               "May", "June", "July", "August", "September",
               "October", "November", "December")

daysInMonth = c(31, 28, 31, 30, 31, 30, 31, 
                  31, 30, 31, 30, 31)

cumDays = cumsum(c(1, daysInMonth))
```

```{r}
makePlotRegion = function(xlim, ylim, bgcolor, ylabels, margins, cityName, xtop = TRUE) {
   #This function is to produce a blank plot that has 
   #the proper axes labels, background color, etc.
   #It is to be used for both the top and bottom plot.
  
   #The parameters are
   #xlim is a two element numeric vector used for the two
   #end points of the x axis
   #ylim is the same as xlim, but for the y axis
   #ylabels is a numeric vector of labels for "tick marks"
   #on the y axis
   #We don't need to x labels because they are Month names
   #margins specifies the size of the plot margins (see mar parameter in par)
   #cityName is a character string to use in the title
   #xtop indicates whether the month names are to appear
   #at the top of the plot or the bottom of the plot
  
  #See the assignment for a pdf image of the plot that is
  #produced as a result of calling this function.
  par(bg=bgcolor, mar=margins)
  plot(NULL, xlim = xlim, ylim = ylim, yaxt = "n", xaxt = "n", axes = FALSE)
  if(xtop) {
  	ylab = sapply(ylabels, function(x) paste0(x, "°"))
  	axis(side = 2, at = ylabels, labels = ylab, tick = TRUE, lty = "solid", lwd = 2, col = "darkgray", col.ticks = "#E2DED5", las = 1, hadj = 1)
  	axis(side = 4, at = ylabels, labels = ylab, tick = TRUE, lty = "solid", lwd = 2, col = "darkgray", col.ticks = "#E2DED5", las = 1, hadj = 1, mgp = c(0,2,0))
  } else {
  	axis(side = 2, at = ylabels, tick = TRUE, lty = "solid", lwd = 2, col = "darkgray", col.ticks = "#E2DED5", las = 1, hadj = 1)
  	axis(side = 4, at = ylabels, tick = TRUE, lty = "solid", lwd = 2, col = "darkgray", col.ticks = "#E2DED5", las = 1, hadj = 1, mgp = c(0,1,0))
  }
  
  title(main=cityName, adj = 0)
  side = 1 + 2 * xtop
  axis(side, at = cumDays[-13] + 15, tick = FALSE, labels = monthNames)
}
```

```{r}

drawTempRegion = function(day, high, low, col){
   #This plot will produce 365 rectangles, one for each day
   #It will be used for the record temps, normal temps, and 
   #observed temps
  rect(xleft = c(0:364), xright = c(1:365), ybottom = low, ytop = high, border = NA, col = col)
   #day - a numeric vector of 365 dates
   #high - a numeric vector of 365 high temperatures
   #low - a numeric vector of 365 low temperatures
   #col - color to fill the rectangles
  
}

addGrid = function(location, col, ltype, vertical = TRUE) {
   #This function adds a set of parallel grid lines
   #It will be used to place vertical and horizontal lines
   #on both temp and precip plots
  if(vertical) {
  	location = location[-1]
  	abline(v = location, lty=ltype, col=col)
  } else {
  	abline(h = location, lty=ltype, col=col)
  }
   #location is a numeric vector of locations for the lines
   #col - the color to make the lines
   #ltype - the type of line to make
   #vertical - indicates whether the lines are vertical or horizontal
}

monthPrecip = function(day, dailyprecip, normal, first = FALSE){
   #This function adds one month's precipitation to the 
   #precipitation plot.
   #It will be called 12 times, once for each month
   #It creates the cumulative precipitation curve,
   #fills the area below with color, add the total
   #precipitation for the month, and adds a reference
   #line and text for the normal value for the month
  points(x = day, y = dailyprecip, col = "#2D5667", type = "l", lwd = 4)
  polygon(x = c(day, max(day), day[1]), y = c(dailyprecip, 0, 0), col = "#C2BCA4", border = NA)
  points(x = c(day[1], max(day)), y = rep(normal, 2), type = "l", col = "#6A848B",lwd = 2)
  if(first) {
  	text(x = day[1] + 1, y = normal[1] + 0.5, labels = paste0("NORMAL\n", normal), cex = .5, adj = 0)
  	text(x = max(day) - 2, y = max(dailyprecip) + 0.05, labels = paste0("ACTUAL  ", max(dailyprecip)), cex = .5, adj = 1)
  } else {
  	text(x = day[1] + 1, y = normal[1] + 0.25, labels = normal, cex = .5, adj = 0)
  	text(x = max(day) - 0.5, y = max(dailyprecip) + 0.2, labels = max(dailyprecip), cex = .5, adj = 1)
  }
   #day a numeric vector of dates for the month
   #dailyprecip a numeric vector of precipitation recorded
   #for the month (any NAs can be set to 0)
   #normal a single value, which is the normal total precip
   #for the month
   #first a boolean value marking the January precip plot
  
}

finalPlot = function(temp, precip){
  # The purpose of this function is to create the whole plot
  # Include here all of the set up that you need for
  # calling each of the above functions.
  # temp is the data frame sfoWeather or laxWeather
  # precip is the data frame sfoMonthlyPrecip or laxMonthlyPrecip

  
  #Here are some vectors that you might find handy
  normPrecip = as.numeric(as.character(precip$normal))
  #Fill in the various stages with your code
 
  
   #Add any additional variables that you will need here

   #Set up the graphics device to plot to pdf and layout
   #the two plots on one canvas

  pdf("LA_PRECIP", width = 14, height = 7.5)
  layout(matrix(c(1,2), nrow = 2, ncol = 1, byrow=TRUE), height = c(5,2.5))
  
   #Call makePlotRegion to create the plotting region
   #for the temperature plot
  ytix = seq(20, 110, by = 10)
  makePlotRegion(c(14,352), c(23,110), "#E2DED5", ytix, c(0, 3, 4, 2.5), "Los Angeles Weather in 2013\n", TRUE)
  
   #Call drawTempRegion 3 times to add the rectangles for
   #the record, normal, and observed temps
  temp = temp[!is.na(temp$High) & !is.na(temp$PrecipYr), ]

  drawTempRegion(c(1:365), temp$RecordLow, temp$RecordHigh, "#C0B9A6")
  drawTempRegion(c(1:365), temp$NormalLow, temp$NormalHigh, "#9E9282")
  drawTempRegion(c(1:365), temp$Low,temp$High,"#5e3b49")
  
  #Call addGrid to add the grid lines to the plot
  addGrid(cumDays,'black', 3)
  addGrid(seq(20, 120, by = 10), col = "#E2DED5", lty = 1, FALSE)
  #Add the markers for the record breaking days
  recHighs = which(temp$High >= temp$RecordHigh)
  recLows = which(temp$Low <= temp$RecordLow)

  sapply(recHighs, function(x) {
  	segments(x0 = x - 0.5, y0 = temp$High[x], x1 = x - 0.5, y1 = temp$High[x] + 10, lwd = 0.5)
  	if(temp$High[x] == temp$RecordHigh[x]) {
  		text(x = x, y = temp$High[x] + 9, labels = paste0("TIED RECORD\nHIGH: ", temp$High[x], "°"), cex = .5, adj = 0)
  	} else {
  		text(x = x, y = temp$High[x] + 9, labels = paste0("RECORD HIGH: ", temp$High[x], "°"), cex = .5, adj = 0)
  	}
  })

  #Ugly fix for adjacent record lows
  for(i in 1:length(recLows)){
  	low = recLows[i]
  	segments(x0 = low - 0.5, y0 = temp$Low[low], x1 = low - 0.5, y1 = temp$Low[low] - 10 - ifelse(i < length(recLows) & i < 3 & (abs(low - recLows[i + 1]) <= 2), 3 * i, 0), lwd = 0.5)
  	if(temp$Low[low] == temp$RecordLow[low]) {
  		text(x = low, y = temp$Low[low] - 9 - ifelse(i < length(recLows) & i < 3 & (abs(low - recLows[i + 1]) <= 2), 3 * i, 0), labels = paste0("TIED RECORD\nLOW: ", temp$Low[low], "°"), cex = .5, adj = 0)
  	} else {
  		text(x = low, y = temp$Low[low] - 9 - ifelse(i < length(recLows) & i < 3 & (abs(low - recLows[i + 1]) <= 2), 3 * i, 0), labels = paste0("RECORD LOW: ", temp$Low[low], "°"), cex = .5, adj = 0)
  	}
  }
  
  #Add the titles
  text(x = 20, y = 110, labels = "Temperature", cex = 1.1, col = "black", font=2)
  text(x = 55.5, y = 103, labels = "Bars represent range between the daily high and low.", cex = .9, col = "black")
  text(x = 54.75, y = 99, labels = "Average daily low temperature for the year was 54.0°,", cex = .9, col = "black")
  text(x = 39.5, y = 95, labels = "and the average daily high was 67.9°.", cex = .9, col = "black")
  
  rect(xleft = 181, xright = 183, ytop = 47, ybottom = 25, col = "#C0B9A6", border = NA)
  rect(xleft = 181, xright = 183, ytop = 42, ybottom = 32, col = "#9E9282", border = NA)
  rect(xleft = 181, xright = 182.25, ytop = 43, ybottom = 35, col = "#5e3b49", border = NA)

  text(x = 171, y = 47, labels = "RECORD HIGH", cex = .4, col = "black")
  text(x = 171, y = 26, labels = "RECORD LOW", cex = .4, col = "black")
  text(x = 168, y = 37, labels = "NORMAL RANGE", cex = .4, col = "black")
  text(x = 190, y = 43, labels = "-ACTUAL HIGH", cex = .4, col = "black")
  text(x = 190, y = 35, labels = "-ACTUAL LOW", cex = .4, col = "black")

  segments(x0 = 177, y0 = 32, x1 = 177, y1 = 42, lwd = 0.5)
  segments(x0 = 177, y0 = 32, x1 = 178, y1 = 32, lwd = 0.5)
  segments(x0 = 177, y0 = 42, x1 = 178, y1 = 42, lwd = 0.5)
  
  
  
  #Call makePlotRegion to create the plotting region
  #for the precipitation plot
  makePlotRegion(c(14, 352), c(0, 5), "#E2DED5", seq(0, 4, by = 1), c(3, 3, 1, 2.5), cityName = "", FALSE)
  
   #Call monthPrecip 12 times to create each months 
   #cumulative precipitation plot. To do this use 
   #sapply(1:12, function(m) {
   #           code
   #             monthPrecip(XXXX)
   #           }) 
   #the anonymous function calls monthPrecip with the 
   #appropriate arguments
  sapply(1:12, function(m) {
  	if(m == 1) {
  		monthPrecip(cumDays[m] + temp$Day[temp$Month == m], cumsum(temp$Precip[temp$Month == m]), normPrecip[m], TRUE)
  	} else {
  		monthPrecip(cumDays[m] + temp$Day[temp$Month == m], cumsum(temp$Precip[temp$Month == m]), normPrecip[m])
  	}
  })

  #To fix issue with normal line not aligning properly
  cumDaysFix = cumsum(c(33, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31))

  #Call addGrid to add the grid lines to the plot
  addGrid(location = seq(0, 5, by = 1), col = "#E2DED5", ltype = "solid", vertical = FALSE)
  segments(x0 = cumDaysFix , y0 = 0, y1 = 4, col = "black", lty="dotted")
  
  #Add the titles
  text(x = 20, y = 4.25, labels = "Precipitation", cex= 1.1, font = 2, pos = 3)
  text(x = 155, y = 4.25, labels = "Cumulative monthly precipitation in inches compared with normal monthly precipitation. Total precipitation in LA was 9.87.", cex= .9, pos=3)

  #Close the pdf device dev.off()
  dev.off()
}

finalPlot(temp = laxWeather, precip = laxMonthlyPrecip)
```

#Part II The Data Life Cycle:

Explain how your plot creation from this homework is described (or not described) by the lifecycle of data.

1. Acquire
While this homework (excluding the extra credit) did not require a rigorous method of data acquisition such as scraping or conducting an experiment, retrieving the data from the NOAA website falls under the "acquire" category of the lifecycle of data.

2. Clean
There were several NA values in Los Angeles' high temperatures as well as the precipitation years which needed to be removed in order to plot the data. Skipping this step would have resulted in large gaps in the plot which would require further explanation. Additionally, the documentation in the code itself, particularly the explanation of function parameters, serve as metadata or annotations for potential future reference.

3. Use/Reuse
I used the cleaned data to create the temperature and precipitation plots which would be a vizualization under the Use/Reuse category.

4. Publish
I uploaded my code to Moodle for graders to reference and/or run, and this written document will serve as an additional outline of the steps I took to turn the raw data into the final visualization.

5. Preserve/Destroy
While I haven't released my code public, I am storing the documents from this assignment on my computer for future reference or work. Some of the functionality can likely be abstracted and generalized further to work on different sets of data. (For the purposes of this assignment I hard-coded most of the graphing parameters, which restricts the flexibility of the code) 