---
title: "HW3"
author: Artur Sak
date: "September 20, 2016"
output: html_document
---

##Q1 (5 pts)
Given a function below,
```{r}

myfunc <- function(z) return(c(z,z^2, z^3%/%2))
```
(1) Examine the following code, and briefly explain what it is doing.

```{r}

y = 2:8
matrix(myfunc(y),ncol=3)
```

This code creates a $3\times7$ matrix consisting of values between 2 and 8 in the first column, their squares in the second, and their cubes (divided by 2 and rounded down to the nearest integer) in the third.

(2) Simplify the code in (1) using one of the "apply" functions and save the result as m.
```{r}

m = t(sapply(y, myfunc))
print(m)
```

(3) Find the row product of m.
```{r}

print(apply(m,1,prod))
```
(4) Find the column sum of m in two ways.
```{r}

print(apply(m,2,sum))
print(colSums(m))
```

(5) Could you divide all the values by 2 in two ways?
```{r}

print(apply(m,2,function(x) x/2))
print(m/2)
```
Yes.

##Q2 (8 pts)
Create a list with 2 elements as follows:
```{r}

l <- list(a = 1:10, b = 11:20)
```
(1) What is the product of the values in each element?
```{r}

print(lapply(l, prod))
```
(2) What is the (sample) variance of the values in each element?

```{r}

print(lapply(l, var))
```
(3) What type of object is returned if you use lapply? sapply? Show your R code that finds these answers.
```{r}

class(lapply(l,var))
class(sapply(l,var))
```

Now create the following list:
```{r}

l.2 <- list(c = c(21:30), d = c(31:40))
```
(4) What is the sum of the corresponding elements of l and l.2, using one function call?

```{r}

print(mapply(sum, l$a,l$b,l.2$c,l.2$d))
```
(5) Take the log of each element in the list l:

```{r}

print(lapply(l, log))
```
(6) First change l and l.2 into matrices, make each element in the list as column,

```{r}

l = sapply(l, cbind)
l.2 = do.call(cbind,l.2)
```

Then, form a list named mylist using l,l.2 and m (from Q1) (in this order).

```{r}

mylist = list(l, l.2, m)
print(mylist)
```

Then, select the first column of each elements in mylist in one function call (hint '[' is the select operator).

```{r}

print(lapply(mylist, '[', ,1))
```

##Q3 (3 pts)
Let's load our friend family data again.
```{r}

load(url("http://courseweb.lis.illinois.edu/~jguo24/family.rda"))
```

(1) Find the mean bmi by gender in one function call.

```{r}

print(tapply(family$bmi, family$gender, mean))
```
(2) Could you get a vector of what the type of variables the dataset is made of？

```{r}

print(sapply(family, class))
```
Yes.

(3) Could you sort the firstName in height descending order?

```{r}

print(family[order(-family$height), c("firstName","height")])
```

Yes.

##Q4 (2 pts)
There is a famous dataset in R called "iris." It should already be loaded in R for you. If you type in ?iris you can see some documentation. Familiarize yourself with this dataset.
```{r, eval=FALSE}

?iris
```

(1) Find the mean petal length by species.

```{r}

print(tapply(iris$Petal.Length, iris$Species, mean))
```

(2) Now obtain the sum of the first 4 variables, by species, but using only one function call.

```{r}
print(aggregate(.~Species, iris, sum))
```

##Q5 (2 pts)
Below are two statements, their results have different structure,

```{r,eval=FALSE}

lapply(1:4, function(x) x^3)
sapply(1:4, function(x) x^3)
```
Could you change one of them to make the two statements return the same results (type of object)?

Yes either pass in 'simplify=FALSE' as a parameter to sapply() or wrap the lapply() in unlist().

```{r}

print(unlist(lapply(1:4, function(x) x^3)))
print(sapply(1:4, function(x) x^3))
print(sapply(1:4, function(x) x^3, simplify=F))
```

##Q6. (5 pts) 

Using the family data, fit a linear regression model to predict weight from height. Place your code and output (the model) below.

```{r}

model = lm(family$weight~family$height)
print(summary(model))
```

How do you interpret this model?

* If height increases by 1 inch, weight is expected to increase by $\approx$ 9.15lbs. 
* The predictive model appears to be a good fit with p-values on the coefficients as well as the overall F-statistic being $<$ 0.05
* The variables appear to have a strong positive correlation ($\approx$ 86%).

Create a scatterplot of height vs weight. Add the linear regression line you found above.

There appears to be a strong positive relationship between height and weight in the family data provided.

```{r}
library(ggplot2)
ggplot(family, aes(x=height, y=weight)) + geom_point(shape=19) + geom_smooth(method=lm) + ggtitle("Family Height vs. Weight") + labs(x="Height", y="Weight")
```
